Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-06-30 paper/tci.tex
Included new story Fig.1

Our work is based on longitudinal data acquisition scenarios where we wish to study new changes that evolve within an object over time, such as in repeated scanning for disease monitoring, or in tomography-guided surgical procedures. While this is easily feasible when measurements are acquired from a large number of projection angles (referred to as `views' henceforth), it is challenging when the number of views is limited. If the goal is to track the changes while simultaneously reducing sub-sampling artefacts, we propose (1) acquiring measurements from a \emph{small} number of views and using a global \emph{`unselective'} prior based reconstruction. If the goal is to observe \emph{details} of new changes, we propose (2) acquiring measurements from a \emph{moderate} number of views and using a more involved reconstruction routine. We show that in the latter case, a `selective' technique is necessary in order to prevent the prior from adversely affecting the reconstruction of new structures that are absent in any of the earlier scans. The reconstruction of new regions is safeguarded from the bias of the prior by computing regional weights that moderate the local influence of the priors. We are thus able to effectively reconstruct both the old and the new structures in the test. We have tested the efficacy of our method on synthetic as well as real volume data. The results demonstrate the use of both unselective and selective priors in different scenarios. Our methods significantly improve the overall quality of the reconstructed data while minimizing the number of measurements needed for imaging in longitudinal studies.
\end{abstract}

@@

In this context, measurements in conventional data acquisition techniques are acquired by sampling the physical object uniformly above the Nyquist rate. However, in the last decade, reconstruction from sub-Nyquist sampled data has been made possible due to methods such as the widely used Compressed Sensing (CS) technique~\cite{Donoho,introCS}, which assumes the data to exhibit properties such as sparsity of the underlying image under certain mathematical transforms such as the Discrete Cosine Transform (DCT) or wavelet transforms. Among the many optimizers used in CS, we used the l1-regularized least squares (l1-ls) form and observed that although this optimizer removes most of the artefacts created due to sub-sampling, its reconstruction is blurred when the number of views is very limited, as shown in Fig.~\ref{fig:diff_methods}. Specifically, the reconstruction in Fig.~\ref{fig:humerus_cs} is obtained by minimizing the following cost using the solver in~\cite{l1ls}:
 \begin{equation}
J_{CS}(\boldsymbol{\theta}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1
\label{Eq:simple_cs}
\end{equation}
Here, $\boldsymbol{x} = \boldsymbol{\Upsilon\theta}$ denotes the reconstructed volume, $\boldsymbol{y}$ its measured tomographic projections, $\boldsymbol{\mathcal{R}}$ the tomographic projection operator such that ideally $\boldsymbol{y} = \boldsymbol{\mathcal{R} x}$, and $\boldsymbol{\theta}$ the sparse coefficients of $\boldsymbol{x}$. $\boldsymbol{\Upsilon}$ is the basis in which $\boldsymbol{x}$ is assumed to be sparse, and is chosen to be DCT here.
In addition to this sparsity prior, when object specific prior information is  used as well, the reconstruction improves significantly. The algorithm presented in~\cite{Essam2015} drew inspiration from CS and solved for the reconstruction from limited views using alternating direction method of multipliers. In other techniques like in~\cite{Polak2017}, reconstruction is performed by estimating sparse gradients in the data. In~\cite{Li2015}, the required number of views is reduced by minimizing the total variation in the data. The measurements and an initial estimate of the data guide the iterative routines.

\begin{figure}[!h]
\centering
\fcolorbox{red}{white}{\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template1.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template2.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template3.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template4.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template5.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template6.png}
        \caption{}
     \end{subfigure}}\\
\begin{subfigure}[b]{0.235\linewidth}
        \fcolorbox{yellow}{yellow}{\includegraphics[width=\textwidth]{../images/humerus/global_local/groundTruth.png}}
        \caption{Test}
\end{subfigure}
\quad
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/fbp_humerus.png}
        \caption{FBP}
     \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/reconNoPrior.png}
        \caption{l1-ls}
        \label{fig:humerus_cs}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/global.png}
        \caption{This paper}
     \end{subfigure}
    \caption{\textit{Knowing more about the data helps.} Reconstructions of (g) a test slice  of size $(200,200)$ from the Humerus CT dataset~\cite{humerus} are performed using object-prior [(a)-(f)] and from measurements obtained from only 10 views. (h) Filtered backprojection (FBP) shows streaky sub-sampling artefacts (i) l1-ls reduces the artefacts significantly while slightly blurring details (j) Prior information coupled with l1-ls (CS) greatly improves the reconstruction.}
\label{fig:diff_methods}
\end{figure}

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements.  These prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a particular} previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with the l1-ls form of CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of l1-ls (CS) with the global prior, in two flavors: the vanilla (unselective) and selective global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction -- unselective or selective, is driven by the goal of the procedure.

\begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{../images/prior_cmb.png}
        \caption{Overview of our work. The choice of the number of measurement views and the type of reconstruction is driven by the goal in the application under consideration.  When our goal is relatively simple, such as tracking the location of new changes while simultaneously reducing sub-sampling artefacts, we propose (a) acquiring measurements from a small number of views (`few-view' imaging) and using unselective prior based reconstruction. When our goal becomes more ambitious, such as observing details of the new changes while simultaneously reducing sub-sampling artefacts, we propose (b) acquiring measurements from a slightly higher number of views (`moderate-view' imaging) and using selective prior-based reconstruction. In either case, the number of views is lower than what is conventionally used.}
 \label{fig:prior_overview}
\end{figure}

\begin{enumerate}
\item Initially, when our goal is to only track the location of the probe while simultaneously reducing sub-sampling artefacts, we advocate capturing measurements from a small number of views (`few-view' imaging) and using unselective prior based reconstruction. 
\item Later, when the probe is proximal to or touches the tumor, our goal is to observe details of tumor ablation while simultaneously reducing sub-sampling artefacts. In this case, we propose to capture measurements from a slightly higher number of views (`moderate-view' imaging) to acquire more information, and use a selective prior-based reconstruction. The selective technique moderates the effect of the prior in the reconstruction of new changes in the object being scanned.
\end{enumerate}

@@

This paper is organized as follows: Section~\ref{sec:tmh} demonstrates the utility of both the unselective and selective methods on a longitudinal medical dataset. Section~\ref{sec:unselective_global_prior} describes the construction of the unselective global eigenspace prior. In addition, we recap the advantages of global prior over dictionary priors. Section~\ref{sec:selective_prior} describes how the unselective prior needs to be modified when accurate details of new changes are to be observed. A selective technique offers a solution. Results validated on both synthetic and real 3D biological datasets are shown. Finally, we conclude with key inferences that can be drawn from our work in Section~\ref{sec:conclusions}.

@@

\section{Application: Reconstruction for CT-guided radio-ablation study}
\label{sec:tmh}
Before diving into the details of the unselective and the selective global prior methods, we first show how both the techniques can be applied to our advantage in a real-life medical longitudinal study. Longitudinal studies here refers to the acquisition of sequential CT scans of the same subject to track time-evolving changes within the subject's interior. Here, our data~\footnote{Source: Tata Memorial Centre~\cite{tmh}, Parel, Mumbai.  This is the national comprehensive centre for the prevention, treatment, education and research in cancer, and is recognized as one of the leading cancer centres in India.} consists of successive scans of the liver taken during a radio-frequency ablation procedure described in the previous section. In such a procedure, the physician inserts a thin needle-like probe into the organ~\cite{rfa}. Once the needle hits the tumor, a high-frequency current is passed through the tip of the probe and this burns the malignant tumor. Throughout this process, multiple CT scans help the physician to track the position of the needle and check the changes within. In this context, we classify the goal of any of our reconstruction techniques into two categories:

@@

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) Filtered backprojection and no prior resulting in streaks, SSIM $= 0.48$ (c) l1-ls resulting in blurred bone structures, SSIM $=0.35$ and (d) unselective global prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: selective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) l1-ls with SSIM $=0.46$ and (d) unselective global prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) selective prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

@@

The needle has a very high attenuation coefficient when compared to that of the organs. Hence, the needle can be tracked by acquiring measurements from a very small number of views. We use the unselective global prior reconstruction here to reduce the artefacts due to sub-sampling. The unselective method is fast and sufficient to track the position of the needle.  Once the needle reaches the site of the tumor, we propose changing the imaging protocol to acquire measurements from a moderate number of views. This will enable us to get more information about the new changes. In addition, we then deploy the selective prior method in order to locate the regions of new changes and penalize any dominance of the prior in these regions.  Regardless of the imaging protocol we use (`few' or `moderate'), the number of views is smaller (atleast one-fifth) than the conventional number of views used in a standard hospital setting.

@@

\textbf{Observing details of the ablation:} Next, we choose slices 1-7 as our object-prior and reconstruct slice 8 from 120 views i.e. a somewhat higher number of views this time. Fig.~\ref{fig:RFA2_few_views} shows the reconstructions of slice 8 by different methods. We see that the selective prior reconstruction brings in the advantage of the prior without it adversely affecting the new regions.

%---------------------------------Unselective section---------------------
\section{Method: unselective global prior-based reconstruction}
\label{sec:unselective_global_prior}
Having presented the application, we first review the algorithm~\cite{my_dicta_paper} for a global (unselective) prior-based reconstruction in Sec.~\ref{sec:unselective_global_prior}. This is followed by details of our technique in Sec.~\ref{sec:selective_prior}. Our method estimates the location and magnitude of new changes, and eventually prevents the prior from adversely affecting the reconstruction of new regions in the test. This, we refer to as a \textit{selective} global prior-based reconstruction routine.

@@

 \section{Resuls: Reconstruction by unselective prior}

>>>>

Our work is based on longitudinal data acquisition scenarios where we wish to study new changes that evolve within an object over time, such as in repeated scanning for disease monitoring, or in tomography-guided surgical procedures. While this is easily feasible when measurements are acquired from a large number of projection angles (referred to as `views' henceforth), it is challenging when the number of views is limited. If the goal is to track the changes while simultaneously reducing sub-sampling artefacts, we propose (1) acquiring measurements from a \emph{small} number of views and using a global \emph{`uniform'} prior based reconstruction. If the goal is to observe \emph{details} of new changes, we propose (2) acquiring measurements from a \emph{moderate} number of views and using a more involved reconstruction routine. We show that in the latter case, a `spatially-varying' technique is necessary in order to prevent the prior from adversely affecting the reconstruction of new structures that are absent in any of the earlier scans. The reconstruction of new regions is safeguarded from the bias of the prior by computing regional weights that moderate the local influence of the priors. We are thus able to effectively reconstruct both the old and the new structures in the test. We have tested the efficacy of our method on synthetic as well as real volume data. The results demonstrate the use of both uniform and spatially-varying priors in different scenarios. Our methods significantly improve the overall quality of the reconstructed data while minimizing the number of measurements needed for imaging in longitudinal studies.
\end{abstract}

@@

In this context, measurements in conventional data acquisition techniques are acquired by sampling the physical object uniformly above the Nyquist rate. In such a case when there are sufficient measurements, reconstruction using the conventional filtered backprojection (FBP) suffices, as seen in Fig.~\ref{fig:story1}. However, in the last decade, reconstruction from sub-Nyquist sampled data has been made possible due to methods such as the widely used Compressed Sensing (CS) technique~\cite{Donoho,introCS}, which assumes the data to exhibit properties such as sparsity of the underlying image under certain mathematical transforms such as the Discrete Cosine Transform (DCT) or wavelet transforms. Fig.~\ref{fig:story2} shows how reconstruction using CS routine is better than FBP for sub-Nyquist number of views. Here, among the many optimizers used in CS, we have used the l1-regularized least squares (l1-ls) form and observed that although this optimizer removes most of the artefacts created due to sub-sampling, its reconstruction is blurred when the number of views is very limited, as shown in Fig.~\ref{fig:story2}. Specifically, the reconstruction in Fig.~\ref{fig:story2-CS} and~\ref{fig:story3-CS} is obtained by minimizing the following cost using the solver in~\cite{l1ls}:
 \begin{equation}
J_{CS}(\boldsymbol{\theta}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1
\label{Eq:simple_cs}
\end{equation}
Here, $\boldsymbol{x} = \boldsymbol{\Upsilon\theta}$ denotes the reconstructed volume, $\boldsymbol{y}$ its measured tomographic projections, $\boldsymbol{\mathcal{R}}$ the tomographic projection operator such that ideally $\boldsymbol{y} = \boldsymbol{\mathcal{R} x}$, and $\boldsymbol{\theta}$ the sparse coefficients of $\boldsymbol{x}$. $\boldsymbol{\Upsilon}$ is the basis in which $\boldsymbol{x}$ is assumed to be sparse, and is chosen to be DCT here.
In addition to this sparsity prior, when object specific prior information is  used as well, the reconstruction improves significantly. The algorithm presented in~\cite{Essam2015} drew inspiration from CS and solved for the reconstruction from limited views using alternating direction method of multipliers. In other techniques like in~\cite{Polak2017}, reconstruction is performed by estimating sparse gradients in the data. In~\cite{Li2015}, the required number of views is reduced by minimizing the total variation in the data. The measurements and an initial estimate of the data guide the iterative routines.

\begin{figure}[!h]
\centering
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/testIm.png}
        \caption{Test}
        \label{fig:story1-test}
     \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/600_views/fbp.png}
        \caption{FBP, no prior}
        \label{fig:1000_views_fbp}
    \end{subfigure}
    %\begin{subfigure}[b]{0.23\linewidth}
     %   \includegraphics[width=\textwidth]{../images/story/600_views/CS_lambda0_0.00.png}
     %   \caption{Least squares}
     %\end{subfigure}
    \caption{Reconstruction from large number of views (600 views); (a) 2D test slice of size [260, 260] (b) FBP reconstruction (SSIM = 0.92).}% (c) Least squares reconstruction (SSIM = 0.99).}
\label{fig:story1}
\end{figure}

\begin{figure}[!h]
\centering
    %\begin{subfigure}[b]{0.23\linewidth}
     %   \includegraphics[width=\textwidth]{../images/story/testIm.png}
      %  \caption{Test}
     %\end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/100_views/fbp.png}
        \caption{FBP, no prior}
        \label{fig:1000_views_fbp}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/100_views/CS_lambda0_2.00.png}
        \caption{sparsity prior}
        \label{fig:story2-CS}
     \end{subfigure}
    \caption{Reconstruction of Fig.~\ref{fig:story1-test} from sub-Nyquist number of views (100 views); (a) FBP reconstruction (SSIM = 0.83) (b) l1-ls reconstruction (SSIM = 0.90).}
\label{fig:story2}
\end{figure}

\begin{figure}[!h]
\centering
    %\begin{subfigure}[b]{0.23\linewidth}
     %   \includegraphics[width=\textwidth]{../images/story/testIm.png}
      %  \caption{Test}
     %\end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/40_views/fbp.png}
        \caption{FBP,\\ no prior}
        \label{fig:1000_views_fbp}
    \end{subfigure}
    \quad
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/40_views/CS_lambda0_3.00.png}
        \caption{sparsity \\prior}
        \label{fig:story3-CS}
    \end{subfigure}
    \quad
        \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/story/40_views/plain_pca.png}
        \caption{sparsity \\+ object prior}
     \end{subfigure}
    \caption{Reconstruction of Fig.~\ref{fig:story1-test} from sub-sub-Nyquist number of views (40 views); (a) FBP reconstruction (SSIM = 0.74) (b) l1-ls reconstruction (SSIM = 0.84) (c) prior-based reconstruction (SSIM = 0.85).}
\label{fig:story3}
\end{figure}

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements. As seen in Fig.~\ref{fig:story3}, the use of object prior further improves the reconstruction when the number of views is far below the Nyquist criterion. The prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a particular} previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. %Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with the l1-ls form of CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of l1-ls (CS) with the global prior, in two flavors: the vanilla (uniform) and spatially-varying global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction -- uniform or spatially-varying, is driven by the goal of the procedure.

\begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{../images/prior_cmb.png}
        \caption{Overview of our work. The choice of the number of measurement views and the type of reconstruction is driven by the goal in the application under consideration.  When our goal is relatively simple, such as tracking the location of new changes while simultaneously reducing sub-sampling artefacts, we propose (a) acquiring measurements from a small number of views (`few-view' imaging) and using uniform prior based reconstruction. When our goal becomes more ambitious, such as observing details of the new changes while simultaneously reducing sub-sampling artefacts, we propose (b) acquiring measurements from a slightly higher number of views (`moderate-view' imaging) and using spatially-varying prior-based reconstruction. In either case, the number of views is lower than what is conventionally used.}
 \label{fig:prior_overview}
\end{figure}

\begin{enumerate}
\item Initially, when our goal is to only track the location of the probe while simultaneously reducing sub-sampling artefacts, we advocate capturing measurements from a small number of views (`few-view' imaging) and using uniform prior based reconstruction. 
\item Later, when the probe is proximal to or touches the tumor, our goal is to observe details of tumor ablation while simultaneously reducing sub-sampling artefacts. In this case, we propose to capture measurements from a slightly higher number of views (`moderate-view' imaging) to acquire more information, and use a spatially-varying prior-based reconstruction. The spatially-varying technique moderates the effect of the prior in the reconstruction of new changes in the object being scanned.
\end{enumerate}

@@

This paper is organized as follows: Section~\ref{sec:tmh} demonstrates the utility of both the uniform and spatially-varying methods on a longitudinal medical dataset. Section~\ref{sec:uniform_global_prior} describes the construction of the uniform global eigenspace prior. In addition, we recap the advantages of global prior over dictionary priors. Section~\ref{sec:spatially-varying_prior} describes how the uniform prior needs to be modified when accurate details of new changes are to be observed. A spatially-varying technique offers a solution. Results validated on both synthetic and real 3D biological datasets are shown. Finally, we conclude with key inferences that can be drawn from our work in Section~\ref{sec:conclusions}.

@@

\section{Application: Reconstruction for CT-guided radio-ablation study}
\label{sec:tmh}
Before diving into the details of the uniform and the spatially-varying global prior methods, we first show how both the techniques can be applied to our advantage in a real-life medical longitudinal study. Longitudinal studies here refers to the acquisition of sequential CT scans of the same subject to track time-evolving changes within the subject's interior. Here, our data~\footnote{Source: Tata Memorial Centre~\cite{tmh}, Parel, Mumbai.  This is the national comprehensive centre for the prevention, treatment, education and research in cancer, and is recognized as one of the leading cancer centres in India.} consists of successive scans of the liver taken during a radio-frequency ablation procedure described in the previous section. In such a procedure, the physician inserts a thin needle-like probe into the organ~\cite{rfa}. Once the needle hits the tumor, a high-frequency current is passed through the tip of the probe and this burns the malignant tumor. Throughout this process, multiple CT scans help the physician to track the position of the needle and check the changes within. In this context, we classify the goal of any of our reconstruction techniques into two categories:

@@

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) Filtered backprojection and no prior resulting in streaks, SSIM $= 0.48$ (c) l1-ls resulting in blurred bone structures, SSIM $=0.35$ and (d) uniform global prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: uniform\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: spatially-varying prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) l1-ls with SSIM $=0.46$ and (d) uniform global prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) spatially-varying prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

@@

The needle has a very high attenuation coefficient when compared to that of the organs. Hence, the needle can be tracked by acquiring measurements from a very small number of views. We use the uniform global prior reconstruction here to reduce the artefacts due to sub-sampling. The uniform method is fast and sufficient to track the position of the needle.  Once the needle reaches the site of the tumor, we propose changing the imaging protocol to acquire measurements from a moderate number of views. This will enable us to get more information about the new changes. In addition, we then deploy the spatially-varying prior method in order to locate the regions of new changes and penalize any dominance of the prior in these regions.  Regardless of the imaging protocol we use (`few' or `moderate'), the number of views is smaller (atleast one-fifth) than the conventional number of views used in a standard hospital setting.

@@

\textbf{Observing details of the ablation:} Next, we choose slices 1-7 as our object-prior and reconstruct slice 8 from 120 views i.e. a somewhat higher number of views this time. Fig.~\ref{fig:RFA2_few_views} shows the reconstructions of slice 8 by different methods. We see that the spatially-varying prior reconstruction brings in the advantage of the prior without it adversely affecting the new regions.

%---------------------------------Uniform section---------------------
\section{Method: uniform prior-based reconstruction}
\label{sec:uniform_global_prior}
Having presented the application, we first review the algorithm~\cite{my_dicta_paper} for a global (uniform) prior-based reconstruction in Sec.~\ref{sec:uniform_global_prior}. This is followed by details of our technique in Sec.~\ref{sec:spatially-varying_prior}. Our method estimates the location and magnitude of new changes, and eventually prevents the prior from adversely affecting the reconstruction of new regions in the test. This, we refer to as a \textit{spatially-varying} global prior-based reconstruction routine.

@@

 \section{Results: Reconstruction by uniform prior}

 The proposed method has been validated on new\footnote {These and our code will be made available to the community.} scans of biological specimens in a longitudinal setting. These datasets in the form of \emph{raw cone-beam projection measurements} were acquired from a lab at the Australian National University (ANU). %We emphasize that in most of the literature on tomographic reconstruction, the results are shown on reconstruction from projections \emph{simulated} from 3D volumes. This is because most CT scanners do not reveal the raw projections, and instead output only the full reconstructed volumes. Moreover, the process of conversion from the projections to the full volumes is proprietary. \emph{Departing from this, we demonstrate reconstruction results from raw projection data.}
In all figures in this section, `uniform prior' refers to optimizing Eq.~\ref{eq:spatially-varying_prior} with $\boldsymbol{W}(x,y,z)=1$.
%----------------------------------------------------------
%-------------------------------------------------
\subsubsection{\textbf{Okra dataset}}
\label{Sec:okra_uniform}
Our first dataset is that of an Okra specimen consisting of its five scans (Fig.~\ref{fig:object-prior_test_okra}). The measurements consisted of real cone-beam projections from 450 views, each of size $336\times156$. The corresponding size of the reconstructed volume is $338\times338\times123$. Prior to the first scan, two holes were drilled on the surface of the specimen. This was followed by four scans, each after introducing one new cut. The specimen was kept in the same position throughout the acquisitions. In cases where such an alignment is not present, all the template volumes must be pre-aligned before computing the eigenspace. The test must be registered to the object-prior after its preliminary pilot reconstruction. The ground truth consists of FDK reconstructed volumes from the the full set of 450 view projections. The test volume was reconstructed from a partial set of 45 projections, i.e. $10\%$ of the projection views from which ground truth was reconstructed. The selected 3D ground truth of template volumes, the test volume as well as the 3D reconstructions can be seen in~\cite{supp_paper}. One of the slices of the reconstructed volumes is shown in Fig.~\ref{fig:okra_3D_results}. The red and green 3D RoI in the video and images show the regions where new changes are present. Based on the potato and the okra experiments, we see that our method is able to \emph{discover both} the presence of a new structure, as well as the absence of a structure.
